Preparing your manuscript

### 2025-01-30 article/article.qmd
Figure 1, version 1

library(metR)
library(ggplot2)
library(data.table)
library(lubridate)
library(here)

@@

```{r}
readr::read_rds(here("data/cmip_file_list_scenario.rds")) |> 
  _[nominal_resolution == "100 km" & source_id %in% models_in_gadi, 
    .(source_id, institution_id, nominal_resolution)] |> 
  unique(by = "source_id") |> 
  knitr::kable(col.names = c("Model", "Institution", "Original resolution"),
               caption = "List of models used for the analysis.") |> 
  kableExtra::kable_classic()
```

@@

>>>>

library(metR)
library(ggplot2)
library(data.table)
library(lubridate)
library(here)
library(cowplot)

@@

```{r}
#| label: model-table

readr::read_rds(here("data/cmip_file_list_scenario.rds")) |> 
  _[nominal_resolution == "100 km" & source_id %in% models_in_gadi, 
    .(source_id, institution_id, nominal_resolution)] |> 
  unique(by = "source_id") |> 
  knitr::kable(col.names = c("Model", "Institution", "Original resolution"),
               caption = "List of models used for the analysis.") |> 
  kableExtra::kable_classic()
```

@@

(ref:trend) a) Multimodel change in drops in temperature frequency (%, SSP5-8.5 2018-2100 minus Historical 1980-2000) and zonal mean change over sea b) and land c), each grey line represent the models listed in Table 1 and in black, the multimodel mean.

```{r}
#| label: trend
#| fig-width: 10
#| fig-height: 5
#| fig.cap: (ref:trend)

data <- purrr::map(Sys.glob(here("data/eke_mean/*hist*mean.nc")), function(f) {

  message(f)
  meta <- unglue::unglue(basename(f), patterns = c("{var}_{model}_{scenario}_mean.nc", "{var}_{model}_{scenario}_mean.nc"))

  if (meta[[1]][["var"]] == "eke") {
    var <- "ua"
  } else {
    var <- "tasmax"
  }
  file_scenario <- here(paste0("data/eke_mean/", meta[[1]][["var"]], "_", meta[[1]][["model"]], "_ssp585_mean.nc"))

  ReadNetCDF(f, vars = c(hist = var)) |> 
    _[, let(ssp585 = ReadNetCDF(file_scenario, vars = c(ssp585 = var), out = "vector")[[1]])] |> 
    _[, let(model = meta[[1]][["model"]],
            var = meta[[1]][["var"]],
            change = ssp585 - hist,
            lat = round(lat, digits = 8),
            plev = NULL)] |> 
    _[]

}) |> rbindlist()

map_trend <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "change") |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lon, lat)] |> 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_contour_fill(aes(z = mean_dt/365*100, fill = after_stat(level)),
                    breaks = c(-Inf, seq(-3, 3, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorbar(barheight = 0.5),
                                   labels = function(x) JumpBy(x, 2, fill = "")) +
  geom_sf(data = global_map, inherit.aes = FALSE, fill = NA, linewidth = 0.2) +
  # scale_y_latitude(labels = NULL) +
  coord_sf(expand = FALSE) +
  labs(x = NULL, y = NULL, fill = "%") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
        legend.key.width = grid::unit(1, 'null'),
        # axis.title.x=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), plot.margin = unit(c(0, 0 , 0 , 0), units = "mm"))

mean_trend <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "change") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lat, land, model)] |> 
  ggplot(aes(mean_dt, lat)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_vline(xintercept = 0, color = "grey80") +
  geom_path(aes(group = model), color = "grey70", alpha = 0.7) +
  stat_summary(orientation = "y", geom = "line") +
  scale_y_latitude(breaks = seq(-80, 80 , 20)) +
  facet_wrap(~land) +
  tagger::tag_facets(tag_pool = c("b", "c")) +
  labs(x = "Change (%)") +
  coord_fixed(ratio = 1/2.8) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text.x = element_text(size = 12),
        plot.margin = unit(c(0, 0 , 1.9, -0.40), units = "cm"),
        tagger.panel.tag.text = element_text(size = 12),
         tagger.panel.tag.background = element_rect(fill = "white", color = "white"))

plot_grid(map_trend, mean_trend, labels = c('a)', ''), 
          label_x = 0, label_y = 0.9,
          label_fontface = "plain", 
          label_colour = "white",
          rel_widths = c(1.75, 1))

```
